---
title: "Understat Primera Vuelta"
author: "Eric Morris"
date: "1/21/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(viridis)
library(readxl)
library(ggrepel)

knitr::opts_chunk$set(
  fig.align = "center",
  dpi = 800
)
```

## La Liga Primera Vuelta (Half Season) Metrics from Understat

All data was pulled manually from Understat (scraping was difficult due to UI). Statistics were filtered for players with a minimum of 500 minutes played. Keepers were filtered out.

```{r Import}
understat = 
  read_excel("./data/understat_primeravuelta.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(x_g, into = c("xg", "trash"), sep = "\\+") %>% 
  separate(xg, into = c("xg", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  separate(n_px_g, into = c("npxg", "trash"), sep = "\\+") %>% 
  separate(npxg, into = c("npxg", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  separate(x_a, into = c("xa", "trash"), sep = "\\+") %>% 
  separate(xa, into = c("xa", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  mutate(xg90_plus_xa90 = x_g90 + x_a90,
         npxg90_plus_xa90 = n_px_g90 + x_a90,
         xg = as.numeric(xg),
         xa = as.numeric(xa))
```


```{r ImportLeague}
team_understat = 
  read_excel("./data/understat_laliga_team.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  select(-g, -apps) %>% 
  separate(team_x_g, into = c("team_xg", "trash"), sep = "\\+") %>% 
  separate(team_xg, into = c("team_xg", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  separate(team_n_px_g, into = c("team_npxg", "trash"), sep = "\\+") %>% 
  separate(team_npxg, into = c("team_npxg", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  separate(team_x_ga, into = c("team_xga", "trash"), sep = "\\+") %>% 
  separate(team_xga, into = c("team_xga", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  separate(x_pts, into = c("x_pts", "trash"), sep = "\\+") %>% 
  separate(x_pts, into = c("x_pts", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2)
```


```{r LaLigaMerge and Chart}
laliga_merge = 
  left_join(understat, team_understat, by = "team") %>% 
  select(-position, -w, -d, -l, -ga, -pts, -x_pts) %>% 
  mutate(team_xg = as.numeric(team_xg),
         team_npxg = as.numeric(team_npxg),
         team_xga = as.numeric(team_xga),
         npxg = as.numeric(npxg),
         player_contribution = (npxg / team_npxg) * 100)

##charting player %

quantile(understat$npxg90_plus_xa90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

quantile(laliga_merge$player_contribution, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

laliga_merge %>% 
  ggplot(aes(x = npxg90_plus_xa90, y = player_contribution, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(npxg90_plus_xa90 >= 0.5, player, ifelse(player_contribution >= 17.795, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta - Offensive Value", 
       y = "Players NPxG / Team Total NPxG (%)", 
       x = "Non-Penalty Expected Goals plus Expected Assists per 90", 
       color = "Player's Team",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played
       Dashed line represents 50th percentile") + 
  scale_x_continuous(breaks = seq(0, 1.25, 0.05)) + 
  scale_y_continuous(breaks = seq(0, 34, 1)) + 
  geom_hline(yintercept = 3.95, linetype = "dashed", color = "black") + 
  geom_vline(xintercept = 0.17, linetype = "dashed", color = "black") +
  theme_bw() +
  theme(legend.position = "bottom")

no_messi_merge = 
  laliga_merge %>% 
  filter(player != "Lionel Messi")

quantile(no_messi_merge$npxg90_plus_xa90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

quantile(no_messi_merge$player_contribution, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

no_messi_merge %>% 
  ggplot(aes(x = npxg90_plus_xa90, y = player_contribution, color = team, size = npg)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(npxg90_plus_xa90 >= 0.5, player, ifelse(player_contribution >= 17.6519, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta - Offensive Value (Messi Filtered Out)", 
       y = "Players NPxG / Team Total NPxG (%)", 
       x = "Non-Penalty Expected Goals plus Expected Assists per 90", 
       color = "Player's Team",
       size = "Non-Penalty Goals",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played
       Dashed line represents 50th percentile") + 
  scale_x_continuous(breaks = seq(0, 1.25, 0.05)) + 
  scale_y_continuous(breaks = seq(0, 34, 1)) + 
  geom_hline(yintercept = 3.95, linetype = "dashed", color = "black") + 
  geom_vline(xintercept = 0.17, linetype = "dashed", color = "black") +
  theme_bw() +
  theme(legend.position = "bottom")

no_messi_merge %>% 
  ggplot(aes(x = player_contribution, y = npxg90_plus_xa90, color = team, size = npg)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(npxg90_plus_xa90 >= 0.5, player, ifelse(player_contribution >= 17.6519, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta (through 1/14) - Offensive Value (Messi Filtered Out)", 
       y = "Non-Penalty Expected Goals plus Expected Assists per 90", 
       x = "Players NPxG / Team Total NPxG (%)", 
       color = "Player's Team",
       size = "Non-Penalty Goals",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played
       Dashed line represents 50th percentile") + 
  scale_x_continuous(breaks = seq(0, 34, 1)) + 
  scale_y_continuous(breaks = seq(0, 1.25, 0.05)) + 
  geom_hline(yintercept = 0.17, linetype = "dashed", color = "black") + 
  geom_vline(xintercept = 3.95, linetype = "dashed", color = "black") +
  theme_bw() +
  theme(legend.position = "bottom")
```


```{r Charts}
quantile(understat$n_px_g90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

quantile(understat$x_a90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

understat %>% 
  group_by(team) %>% 
  summarize()

understat %>% 
  ggplot(aes(x = n_px_g90, y = x_a90, color = team, size = npg)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(n_px_g90 >= 0.438, player, ifelse(x_a90 >= 0.258, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta Leaders", 
       y = "Expected Assists per 90", 
       x = "Non-Penalty Expected Goals per 90", 
       color = "Player's Team",
       size = "Non-Penalty Goals",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 5% of either category, min. 500 minutes played
       Dashed line represents 50th percentile") + 
  scale_x_continuous(breaks = seq(0, 0.75, 0.05)) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.05)) + 
  geom_hline(yintercept = 0.07, linetype = "dashed", color = "black") + 
  geom_vline(xintercept = 0.08, linetype = "dashed", color = "black") +
  theme_bw() +
  theme(legend.position = "bottom")


understat %>% 
  ggplot(aes(x = n_px_g90, y = x_a90, color = team, size = npg)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(n_px_g90 >= 0.336, player, ifelse(x_a90 >= 0.2, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta Leaders", 
       y = "Expected Assists per 90", 
       x = "Non-Penalty Expected Goals per 90", 
       color = "Player's Team",
       size = "Non-Penalty Goals",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played") + 
  scale_x_continuous(breaks = seq(0, 0.75, 0.05)) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.05)) + 
  theme_bw() +
  theme(legend.position = "bottom")

quantile(no_messi$g, c(0.5, 0.75, 0.80, 0.85, 0.90))

quantile(no_messi$a, c(0.5, 0.75, 0.80, 0.85, 0.90))

understat %>% 
  ggplot(aes(x = g, y = a, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(g >= 4, player, ifelse(a >= 3, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta Leaders", 
       y = "Assists", 
       x = "Goals Scored", 
       color = "Player's Team",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played") + 
  scale_x_continuous(breaks = seq(0, 18, 1)) + 
  scale_y_continuous(breaks = seq(0, 11, 1)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

no_messi %>% 
  ggplot(aes(x = g, y = a, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(g >= 4, player, ifelse(a >= 3, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta Leaders (Messi Filtered)", 
       y = "Assists", 
       x = "Goals Scored", 
       color = "Player's Team",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played") + 
  scale_x_continuous(breaks = seq(0, 15, 1)) + 
  scale_y_continuous(breaks = seq(0, 7, 1)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

quantile(understat$xg, c(0.5, 0.75, 0.80, 0.85, 0.90))

quantile(understat$xa, c(0.5, 0.75, 0.80, 0.85, 0.90))

understat %>% 
  ggplot(aes(x = xg, y = xa, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(xg >= 4.466, player, ifelse(xa >= 2.786, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta Leaders", 
       y = "xA (Total)", 
       x = "xG (Total)", 
       color = "Player's Team",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played") + 
  scale_x_continuous(breaks = seq(0, 15, 0.5)) + 
  scale_y_continuous(breaks = seq(0, 9, 0.5)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

#Messi filtered out

no_messi = understat %>% 
  filter(player != "Lionel Messi")

quantile(no_messi$n_px_g90, c(0.5, 0.75, 0.80, 0.85, 0.90))

quantile(no_messi$x_a90, c(0.5, 0.75, 0.80, 0.85, 0.90))

no_messi %>% 
  ggplot(aes(x = n_px_g90, y = x_a90, color = team, size = npg)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(n_px_g90 >= 0.33, player, ifelse(x_a90 >= 0.2, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta Leaders (Messi Filtered)", 
       y = "Expected Assists per 90", 
       x = "Non-Penalty Expected Goals per 90", 
       color = "Player's Team",
       size = "Non-Penalty Goals",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played") + 
  scale_x_continuous(breaks = seq(0, 0.75, 0.05)) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.05)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

quantile(understat$npg, c(0.5, 0.75, 0.80, 0.85, 0.90))

quantile(understat$npxg90_plus_xa90, c(0.5, 0.75, 0.80, 0.85, 0.90))

understat %>% 
  ggplot(aes(x = npg, y = npxg90_plus_xa90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(npg >= 3.6, player, ifelse(npxg90_plus_xa90 >= 0.5, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Athletic Club" = "red2",
                                "Alaves" = "blue",
                                "Atletico Madrid" = "red",
                                "Barcelona" = "maroon",
                                "Celta Vigo" = "lightblue",
                                "Eibar" = "dodgerblue4",
                                "Espanyol" = "mediumblue",
                                "Getafe" = "royalblue2",
                                "Girona" = "firebrick1",
                                "Leganes" = "dodgerblue",
                                "Levante" = "red4",
                                "Rayo Vallecano" = "gold",
                                "Real Betis" = "green3",
                                "Real Madrid" = "purple1",
                                "Real Sociedad" = "royalblue3",
                                "Real Valladolid" = "mediumpurple",
                                "SD Huesca" = "midnightblue",
                                "Sevilla" = "firebrick3",
                                "Valencia"  = "gray0",
                                "Villarreal" = "yellow3")) + 
  labs(title = "La Liga Primera Vuelta Leaders", 
       y = "Non-Penalty xG + xA per 90", 
       x = "Non-Penalty Goals Scored", 
       color = "Player's Team",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played") + 
  scale_x_continuous(breaks = seq(0, 17, 1)) + 
  scale_y_continuous(breaks = seq(0, 1.25, 0.05)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

```

##EPL Import

All data was pulled manually from Understat (scraping was difficult due to UI). Statistics were filtered for players with a minimum of 500 minutes played. Keepers were filtered out. Data was for period up until 12/28/2018 (end of matchday 19).

```{r EPL Import}
epl_understat = 
  read_excel("./data/understat_epl.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(x_g, into = c("xg", "trash"), sep = "\\+") %>% 
  separate(xg, into = c("xg", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  separate(n_px_g, into = c("npxg", "trash"), sep = "\\+") %>% 
  separate(npxg, into = c("npxg", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  separate(x_a, into = c("xa", "trash"), sep = "\\+") %>% 
  separate(xa, into = c("xa", "trash2"), sep = "\\-") %>% 
  select(-trash, -trash2) %>% 
  mutate(xg90_plus_xa90 = x_g90 + x_a90,
         npxg90_plus_xa90 = n_px_g90 + x_a90)
```

```{r EPL Charts}
quantile(epl_understat$n_px_g90, c(0.5, 0.75, 0.80, 0.85, 0.90))

quantile(epl_understat$x_a90, c(0.5, 0.75, 0.80, 0.85, 0.90))

epl_understat %>% 
  group_by(team) %>% 
  summarize()

epl_understat %>% 
  ggplot(aes(x = n_px_g90, y = x_a90, color = team, size = npg)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(n_px_g90 >= 0.36, player, ifelse(x_a90 >= 0.22, player, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Arsenal" = "firebrick1",
                                "Bournemouth" = "gray0",
                                "Brighton" = "dodgerblue",
                                "Burnley" = "maroon",
                                "Cardiff" = "brown1",     
                                "Chelsea" = "blue",
                                "Crystal Palace" = "purple1",
                                "Everton" = "mediumblue",
                                "Fulham" = "gray48",        
                                "Huddersfield" = "yellow3",   
                                "Leicester" = "cornflowerblue",       
                                "Liverpool" = "red2",       
                                "Manchester City" = "cadetblue2",  
                                "Manchester United" = "firebrick3",
                                "Newcastle United" = "gray12",       
                                "Southampton" = "tomato3",    
                                "Tottenham" = "snow3",       
                                "Watford" = "yellow3",       
                                "West Ham" = "violetred4",      
                                "Wolverhampton Wanderers" = "orange1")) +
  labs(title = "Premier League Primera Vuelta (First Half) Leaders", 
       y = "Expected Assists per 90", 
       x = "Non-Penalty Expected Goals per 90", 
       color = "Player's Team",
       size = "Non-Penalty Goals",
       caption = "Data: understat.com Graph: @ericdavidmorris
       Marked players are in top 10% of either category, min. 500 minutes played") + 
  scale_x_continuous(breaks = seq(0, 0.85, 0.05)) + 
  scale_y_continuous(breaks = seq(0, 0.5, 0.05)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

```

## EPL + La Liga

```{r}
epl_liga = 
  bind_rows(understat, epl_understat)

quantile(epl_liga$n_px_g90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

quantile(epl_liga$x_a90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

quantile(epl_liga$npg, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

epl_liga %>% 
  ggplot(aes(x = n_px_g90, y = x_a90, size = npg, color = league)) +
  geom_point(alpha = 0.5) +
  geom_text_repel(aes(label = ifelse(n_px_g90 >= 0.44, player, ifelse(x_a90 >= 0.27, player, ''))), angle = 30, size = 2) +
  labs(title = "EPL & La Liga Primera Vuelta Leaders", 
       y = "Expected Assists per 90", 
       x = "Non-Penalty Expected Goals per 90", 
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 5% of either category") + 
  scale_x_continuous(breaks = seq(0, 0.825, 0.05)) + 
  scale_y_continuous(breaks = seq(0, 0.6, 0.05)) + 
  theme_bw() +
  scale_color_manual(values = c("La Liga" = "black",
                                "Premier League" = "purple4")) + 
  theme(legend.position = "bottom") 
```

