---
title: "WhoScored Data"
author: "Eric Morris"
date: "12/30/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(viridis)
library(readxl)
library(ggrepel)

knitr::opts_chunk$set(
  fig.align = "center",
  dpi = 800
)
```

## La Liga

All data was pulled manually from WhoScored (scraping was difficult due to UI). Statistics were filtered for players with a WhoScored's automatic minimum appearances filter (players with total appearances greater than the average number of appearances in Primera Division) and then further filtered to only include players with >500 minutes played. Statistics were pulled on 12/29, during La Liga's Winter Break (all teams have played 17/38 matches except Real Madrid with 16). 

```{r Import}
ws_liga_stats = 
  read_excel("./data/WhoScoredLaLigaStats.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(player, into = c("name", "team"), sep = ",") %>%
  mutate(goals_per_90 = (goals / mins) * 90,
         assists_per_90 = (assists / mins) * 90,
         gpa_per_90 = goals_per_90 + assists_per_90,
         shots_kp_game = shots_per_game + key_passes_per_g)

ws_liga_stats %>% 
  group_by(team) %>% 
  summarize()
  
```


```{r La Liga Chart}
quantile(ws_liga_stats$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90))

ws_liga_stats %>% 
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.479, name, ifelse(shots_kp_game >= 3.45, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c(" Athletic Bilbao" = "red2",
                               " Atletico Madrid" = "red",
                               " Barcelona" = "maroon",
                               " Celta Vigo" = "lightblue",
                               " Deportivo Alaves" = "blue",
                               " Eibar" = "dodgerblue4",
                               " Espanyol" = "mediumblue",
                               " Getafe" = "royalblue2",
                               " Girona" = "firebrick1",
                               " Leganes" = "dodgerblue",
                               " Levante" = "red4",
                               " Rayo Vallecano" = "gold",
                               " Real Betis" = "green3",
                               " Real Madrid" = "purple1",
                               " Real Sociedad" = "royalblue3",
                               " Real Valladolid" = "mediumpurple",
                               " SD Huesca" = "midnightblue",
                               " Sevilla" = "firebrick3",
                               " Valencia"  = "gray0",
                               " Villarreal" = "yellow3")) + 
  labs(title = "La Liga Offensive Leaders at Christmas Break", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 8.4, 0.20)) + 
  scale_y_continuous(breaks = seq(0, 1.90, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

#Messi filtered out

no_messi = ws_liga_stats %>% 
  filter(name != "Lionel Messi")

quantile(no_messi$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90))
  
  
ws_liga_stats %>% 
  filter(name != "Lionel Messi") %>%
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.472, name, ifelse(shots_kp_game >= 3.40, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c(" Athletic Bilbao" = "red2",
                               " Atletico Madrid" = "red",
                               " Barcelona" = "maroon",
                               " Celta Vigo" = "lightblue",
                               " Deportivo Alaves" = "blue",
                               " Eibar" = "dodgerblue4",
                               " Espanyol" = "mediumblue",
                               " Getafe" = "royalblue2",
                               " Girona" = "firebrick1",
                               " Leganes" = "dodgerblue",
                               " Levante" = "red4",
                               " Rayo Vallecano" = "gold",
                               " Real Betis" = "green3",
                               " Real Madrid" = "purple1",
                               " Real Sociedad" = "royalblue3",
                               " Real Valladolid" = "mediumpurple",
                               " SD Huesca" = "midnightblue",
                               " Sevilla" = "firebrick3",
                               " Valencia"  = "gray0",
                               " Villarreal" = "yellow3")) + 
  labs(title = "La Liga Offensive Leaders at Christmas Break (Messi Filtered)", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 5.1, 0.10)) + 
  scale_y_continuous(breaks = seq(0, 1.2, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

```


## Serie A

All data was pulled manually from WhoScored (scraping was difficult due to UI). Statistics were filtered for players with a WhoScored's automatic minimum appearances filter (players with total appearances greater than the average number of appearances in Serie A) and then further filtered to only include players with >500 minutes played. Statistics were pulled on 12/30, all teams have played 19 matches. 

```{r Italy Import}
ws_seriea_stats = 
  read_excel("./data/SerieAWhoScoredStats.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(player, into = c("name", "team"), sep = ",") %>%
  mutate(goals_per_90 = (goals / mins) * 90,
         assists_per_90 = (assists / mins) * 90,
         gpa_per_90 = goals_per_90 + assists_per_90,
         shots_kp_game = shots_per_game + key_passes_per_g)

ws_seriea_stats %>% 
  group_by(team) %>% 
  summarize()
  
```


```{r Serie A Chart}
quantile(ws_seriea_stats$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90))
quantile(ws_seriea_stats$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90))

ws_seriea_stats %>% 
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.535, name, ifelse(shots_kp_game >= 4.2, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("AC Milan" = "firebrick2",
                                "Atalanta" = "mediumblue",
                                "Bologna" = "navy",    
                                "Cagliari" = "midnightblue",    
                                "Chievo" = "gold1",        
                                "Empoli" = "coral1",        
                                "Fiorentina" = "purple4",    
                                "Frosinone" = "yellow3",     
                                "Genoa" = "royalblue4",       
                                "Inter" = "royalblue",
                                "Juventus" = "black",      
                                "Lazio" = "lightblue",         
                                "Napoli" = "dodgerblue",
                                "Parma Calcio 1913" = "gray",         
                                "Roma" = "darkorange",        
                                "Sampdoria" = "royalblue2", 
                                "Sassuolo" = "seagreen",
                                "SPAL 2013" = "steelblue3",   
                                "Torino" = "tomato4",
                                "Udinese" = "gray47")) + 
  labs(title = "Serie A Offensive Leaders after 19 matches", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 8.2, 0.20)) + 
  scale_y_continuous(breaks = seq(0, 1.20, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

```


## La Liga + Serie A 

```{r}
italy_spain_merged = 
  bind_rows(ws_liga_stats, ws_seriea_stats)

quantile(italy_spain_merged$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(italy_spain_merged$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(italy_spain_merged$ws_rating, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))

italy_spain_merged %>% 
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = league)) +
  geom_point(alpha = 0.5) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.512, name, ifelse(shots_kp_game >= 3.78, name, ''))), angle = 30, size = 2) +
  labs(title = "La Liga and Serie A Top Creators as of 12/29", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "League",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 8.4, 0.20)) + 
  scale_y_continuous(breaks = seq(0, 1.9, 0.10)) + 
  theme_bw() +
  scale_color_manual(values = c("La Liga" = "firebrick1",
                                "Serie A" = "green4")) + 
  theme(legend.position = "bottom") 

italy_spain_merged %>% 
  ggplot(aes(x = ws_rating, y = shots_kp_game, color = league)) +
  geom_point(alpha = 0.5) +
  geom_text_repel(aes(label = ifelse(shots_kp_game >= 4.44, name, ifelse(ws_rating >= 7.22, name, ''))), angle = 30, size = 2) +
  labs(title = "La Liga and Serie A Top Creators as of 12/29", 
       y = "Shots plus Key Passes per game", 
       x = "WhoScored Rating", 
       color = "League",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 5% of either category") + 
  scale_x_continuous(breaks = seq(0, 8.8, 0.20)) + 
  scale_y_continuous(breaks = seq(0, 8.4, 0.20)) + 
  theme_bw() +
  scale_color_manual(values = c("La Liga" = "firebrick1",
                                "Serie A" = "green4")) + 
  theme(legend.position = "bottom") 
```

## Bundesliga 

All data was pulled manually from WhoScored (scraping was difficult due to UI). Statistics were filtered for players with a WhoScored's automatic minimum appearances filter (players with total appearances greater than the average number of appearances in Serie A) and then further filtered to only include players with >450 minutes played. Statistics were pulled on 12/30, all teams have played 17 matches. 

```{r Germany Import}
ws_bundesliga_stats = 
  read_excel("./data/BundesLigaWhoScoredStats.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(player, into = c("name", "team"), sep = ",") %>%
  mutate(goals_per_90 = (goals / mins) * 90,
         assists_per_90 = (assists / mins) * 90,
         gpa_per_90 = goals_per_90 + assists_per_90,
         shots_kp_game = shots_per_game + key_passes_per_g)

ws_bundesliga_stats %>% 
  group_by(team) %>% 
  summarize()
  
quantile(ws_bundesliga_stats$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(ws_bundesliga_stats$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(ws_bundesliga_stats$ws_rating, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
```

```{r German Chart}
ws_bundesliga_stats %>% 
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.615, name, ifelse(shots_kp_game >= 3.53, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Augsburg" = "dimgray",
                                "Bayer Leverkusen" = "firebrick4",
                                "Bayern Munich" = "firebrick2",
                                "Borussia Dortmund" = "gold1",
                                "Borussia M.Gladbach" = "green3",
                                "Eintracht Frankfurt" = "gray0",
                                "Fortuna Duesseldorf" = "tomato3",  
                                "Freiburg" = "gray25",       
                                "Hannover 96" = "chartreuse4",       
                                "Hertha Berlin" = "dodgerblue2",       
                                "Hoffenheim" = "blue1",     
                                "Mainz 05" = "red1",       
                                "Nuernberg" = "tomato4",
                                "RasenBallsport Leipzig" = "navy",
                                "Schalke 04" = "blue3",        
                                "VfB Stuttgart" = "orangered2",    
                                "Werder Bremen" = "springgreen3", 
                                "Wolfsburg" = "limegreen")) + 
  labs(title = "Bundesliga Offensive Leaders at Christmas Break", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 6.0, 0.10)) + 
  scale_y_continuous(breaks = seq(0, 2.2, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 
```

##Ligue 1

All data was pulled manually from WhoScored (scraping was difficult due to UI). Statistics were filtered for players with a WhoScored's automatic minimum appearances filter (players with total appearances greater than the average number of appearances in Serie A) and then further filtered to only include players with >450 minutes played. Statistics were pulled on 12/30, during the winter break. 

```{r France Import}
ws_ligue_stats = 
  read_excel("./data/Ligue1WhoScoredStats.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(player, into = c("name", "team"), sep = ",") %>%
  mutate(goals_per_90 = (goals / mins) * 90,
         assists_per_90 = (assists / mins) * 90,
         gpa_per_90 = goals_per_90 + assists_per_90,
         shots_kp_game = shots_per_game + key_passes_per_g)

ws_ligue_stats %>% 
  group_by(team) %>% 
  summarize()
  
quantile(ws_ligue_stats$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(ws_ligue_stats$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(ws_ligue_stats$ws_rating, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
```


```{r France Chart}
ws_ligue_stats %>% 
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.541, name, ifelse(shots_kp_game >= 3.89, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Amiens" = "darkgray",
                                "Angers" = "black", 
                                "Bordeaux" = "navy",
                                "Caen" = "dodgerblue3",
                                "Dijon" = "firebrick2", 
                                "Guingamp" = "darkred",
                                "Lille" = "red2",
                                "Lyon" = "blue1",
                                "Marseille" = "darkturquoise",
                                "Monaco" = "goldenrod",
                                "Montpellier" = "darkorange1",
                                "Nantes" = "yellow2",
                                "Nice" = "gray8",
                                "Nimes" = "red3",
                                "Paris Saint Germain" = "midnightblue",
                                "Reims" = "brown1",
                                "Rennes" = "brown4",
                                "Saint Etienne" = "green4",
                                "Strasbourg" = "cadetblue2",
                                "Toulouse" = "mediumorchid4")) + 
  labs(title = "Ligue 1 Offensive Leaders at Christmas Break", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 6.2, 0.10)) + 
  scale_y_continuous(breaks = seq(0, 1.7, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 
```


##EPL 

All data was pulled manually from WhoScored (scraping was difficult due to UI). Statistics were filtered for players with a WhoScored's automatic minimum appearances filter (players with total appearances greater than the average number of appearances in Serie A) and then further filtered to only include players with >500 minutes played. Statistics were pulled on 12/30. 

```{r}
ws_epl_stats = 
  read_excel("./data/EPLWhoScoredStats.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(player, into = c("name", "team"), sep = ",") %>%
  mutate(goals_per_90 = (goals / mins) * 90,
         assists_per_90 = (assists / mins) * 90,
         gpa_per_90 = goals_per_90 + assists_per_90,
         shots_kp_game = shots_per_game + key_passes_per_g)

ws_epl_stats %>% 
  group_by(team) %>% 
  summarize()
  
quantile(ws_epl_stats$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(ws_epl_stats$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
quantile(ws_epl_stats$ws_rating, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95))
```


```{r EPL Chart}
ws_epl_stats %>% 
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.638, name, ifelse(shots_kp_game >= 3.4, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c("Arsenal" = "firebrick1",
                                "Bournemouth" = "gray0",
                                "Brighton" = "dodgerblue",
                                "Burnley" = "maroon",
                                "Cardiff" = "brown1",     
                                "Chelsea" = "blue",
                                "Crystal Palace" = "purple1",
                                "Everton" = "mediumblue",
                                "Fulham" = "gray48",        
                                "Huddersfield" = "yellow3",   
                                "Leicester" = "cornflowerblue",       
                                "Liverpool" = "red2",       
                                "Manchester City" = "cadetblue2",  
                                "Manchester United" = "firebrick3",
                                "Newcastle United" = "gray12",       
                                "Southampton" = "tomato3",    
                                "Tottenham" = "snow3",       
                                "Watford" = "yellow3",       
                                "West Ham" = "violetred4",      
                                "Wolverhampton Wanderers" = "orange1")) + 
  labs(title = "Premier League Offensive Leaders after Matchday 20", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 5.7, 0.10)) + 
  scale_y_continuous(breaks = seq(0, 1.2, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 
```

# top 5

```{r}
top5 = 
  bind_rows(ws_liga_stats, ws_seriea_stats, ws_bundesliga_stats, ws_epl_stats, ws_ligue_stats)

quantile(top5$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95, 0.975))
quantile(top5$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95, 0.975))
quantile(top5$ws_rating, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95, 0.975))
quantile(top5$key_passes_per_g, c(0.5, 0.75, 0.80, 0.85, 0.90, 0.95, 0.975))

top5 %>% 
  ggplot(aes(x = key_passes_per_g, y = gpa_per_90, color = league)) +
  geom_point(alpha = 0.5) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.7594, name, ifelse(key_passes_per_g >= 1.9, name, ''))), angle = 25, size = 1.5) +
  labs(title = "Top 5 European Leagues' Best Creators as of 12/30", 
       y = "Goals plus Assists per 90", 
       x = "Key Passes per game", 
       color = "League",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 5% of either category") + 
  scale_x_continuous(breaks = seq(0, 4.0, 0.10)) + 
  scale_y_continuous(breaks = seq(0, 2.2, 0.10)) + 
  theme_bw() +
  scale_color_manual(values = c("La Liga" = "black",
                                "Serie A" = "green4",
                                "Bundesliga" = "red1",
                                "Premier League" = "purple4",
                                "Ligue 1" = "yellow3")) + 
  theme(legend.position = "bottom") 

top5 %>% 
  ggplot(aes(x = ws_rating, y = shots_kp_game, color = league)) +
  geom_point(alpha = 0.5) +
  geom_text_repel(aes(label = ifelse(ws_rating >= 7.3085, name, '')), angle = 25, size = 1.5) +
  labs(title = "Top 5 European Leagues' Best according to WhoScored as of 12/30", 
       y = "Shots plus Key Passes per game", 
       x = "WhoScored Rating", 
       color = "League",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 5% of WhoScored Rating") + 
  scale_x_continuous(breaks = seq(0, 8.8, 0.20)) + 
  scale_y_continuous(breaks = seq(0, 8.4, 0.20)) + 
  theme_bw() +
  scale_color_manual(values = c("La Liga" = "black",
                                "Serie A" = "green4",
                                "Bundesliga" = "red1",
                                "Premier League" = "purple4",
                                "Ligue 1" = "yellow3")) + 
  theme(legend.position = "bottom") 
```

