---
title: "WhoScored Data"
author: "Eric Morris"
date: "12/30/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(viridis)
library(readxl)
library(ggrepel)

knitr::opts_chunk$set(
  fig.align = "center",
  dpi = 800
)
```

## La Liga

All data was pulled manually from WhoScored (scraping was difficult due to UI). Statistics were filtered for players with a WhoScored's automatic minimum appearances filter (players with total appearances greater than the average number of appearances in Primera Division) and then furthere filtered to only include players with >500 minutes played. Statistics were pulled on 12/29, during La Liga's Winter Break (all teams have played 17/38 matches except Real Madrid with 16). 

```{r Import}
ws_liga_stats = 
  read_excel("./data/WhoScoredLaLigaStats.xlsx", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  separate(player, into = c("name", "team"), sep = ",") %>%
  mutate(goals_per_90 = (goals / mins) * 90,
         assists_per_90 = (assists / mins) * 90,
         gpa_per_90 = goals_per_90 + assists_per_90,
         shots_kp_game = shots_per_game + key_passes_per_g)

ws_liga_stats %>% 
  group_by(team) %>% 
  summarize()
  
```


```{r La Liga Chart}
quantile(ws_liga_stats$shots_kp_game, c(0.5, 0.75, 0.80, 0.85, 0.90))

ws_liga_stats %>% 
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.479, name, ifelse(shots_kp_game >= 3.45, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c(" Athletic Bilbao" = "red2",
                               " Atletico Madrid" = "red",
                               " Barcelona" = "maroon",
                               " Celta Vigo" = "lightblue",
                               " Deportivo Alaves" = "blue",
                               " Eibar" = "dodgerblue4",
                               " Espanyol" = "mediumblue",
                               " Getafe" = "royalblue2",
                               " Girona" = "firebrick1",
                               " Leganes" = "dodgerblue",
                               " Levante" = "red4",
                               " Rayo Vallecano" = "gold",
                               " Real Betis" = "green3",
                               " Real Madrid" = "purple1",
                               " Real Sociedad" = "royalblue3",
                               " Real Valladolid" = "mediumpurple",
                               " SD Huesca" = "midnightblue",
                               " Sevilla" = "firebrick3",
                               " Valencia"  = "gray0",
                               " Villarreal" = "yellow3")) + 
  labs(title = "La Liga Offensive Leaders at Christmas Break", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 8.4, 0.20)) + 
  scale_y_continuous(breaks = seq(0, 1.90, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

#Messi filtered out

no_messi = ws_liga_stats %>% 
  filter(name != "Lionel Messi")

quantile(no_messi$gpa_per_90, c(0.5, 0.75, 0.80, 0.85, 0.90))
  
  
ws_liga_stats %>% 
  filter(name != "Lionel Messi") %>%
  ggplot(aes(x = shots_kp_game, y = gpa_per_90, color = team)) +
  geom_point(alpha = 0.75) +
  geom_text_repel(aes(label = ifelse(gpa_per_90 >= 0.472, name, ifelse(shots_kp_game >= 3.40, name, ''))), angle = 30, size = 2) +
  scale_color_manual(values = c(" Athletic Bilbao" = "red2",
                               " Atletico Madrid" = "red",
                               " Barcelona" = "maroon",
                               " Celta Vigo" = "lightblue",
                               " Deportivo Alaves" = "blue",
                               " Eibar" = "dodgerblue4",
                               " Espanyol" = "mediumblue",
                               " Getafe" = "royalblue2",
                               " Girona" = "firebrick1",
                               " Leganes" = "dodgerblue",
                               " Levante" = "red4",
                               " Rayo Vallecano" = "gold",
                               " Real Betis" = "green3",
                               " Real Madrid" = "purple1",
                               " Real Sociedad" = "royalblue3",
                               " Real Valladolid" = "mediumpurple",
                               " SD Huesca" = "midnightblue",
                               " Sevilla" = "firebrick3",
                               " Valencia"  = "gray0",
                               " Villarreal" = "yellow3")) + 
  labs(title = "La Liga Offensive Leaders at Christmas Break", 
       y = "Goals plus Assists per 90", 
       x = "Shots plus Key Passes per game", 
       color = "Player's Team",
       caption = "Data: WhoScored Graph: @ericdavidmorris
       Marked players are in top 10% of either category") + 
  scale_x_continuous(breaks = seq(0, 5.1, 0.10)) + 
  scale_y_continuous(breaks = seq(0, 1.2, 0.10)) + 
  theme_bw() +
  theme(legend.position = "bottom") 

```



